---
title: "Appendix: R Code for TRAIN Law Analysis"
format: pdf
editor: source
warning: false
---

# Appendix: R Code

## A.1 Data Loading

```{r, message=FALSE, warning=FALSE}
library(data.table)
library(dplyr)
library(readr)
library(ggplot2)
library(scales)
library(broom)

# Load raw FIES volumes (two-volume structure per year)
v1_2012 <- fread("fies_2012_vol.1.csv")
v2_2012 <- fread("fies_2012_vol.2.csv")

v1_2015 <- fread("fies_2015_vol.1.csv")
v2_2015 <- fread("fies_2015_vol.2.csv")

v1_2018 <- fread("fies_2018_vol.1.csv")
v2_2018 <- fread("fies_2018_vol.2.csv")
```

---

## A.2 Merging Volumes and Deduplication Checks

### 2012

```{r}
keys_2012 <- c("W_REGN", "W_OID", "W_SHSN", "W_HCN")

# Check for duplicates in each volume
nrow(v1_2012[, .N, by = keys_2012][N > 1])
nrow(v2_2012[, .N, by = keys_2012][N > 1])

# Merge volumes: drop overlapping non-key columns from vol.2
setkeyv(v1_2012, keys_2012)
setkeyv(v2_2012, keys_2012)
v2_2012_clean <- v2_2012[, !setdiff(intersect(names(v1_2012), names(v2_2012)), keys_2012),
                          with = FALSE]
fies_2012 <- v1_2012[v2_2012_clean]

# Validate merge
nrow(v1_2012); nrow(fies_2012)
fies_2012[, .N, by = keys_2012][N > 1]
all.equal(v1_2012$RFACT, fies_2012$RFACT)
```

### 2015

```{r}
keys_2015 <- c("W_REGN", "W_ID", "W_SHSN", "W_HCN")

nrow(v1_2015[, .N, by = keys_2015][N > 1])
nrow(v2_2015[, .N, by = keys_2015][N > 1])

setkeyv(v1_2015, keys_2015)
setkeyv(v2_2015, keys_2015)
v2_2015_clean <- v2_2015[, !setdiff(intersect(names(v1_2015), names(v2_2015)), keys_2015),
                          with = FALSE]
fies_2015 <- v1_2015[v2_2015_clean]

nrow(v1_2015); nrow(fies_2015)
fies_2015[, .N, by = keys_2015][N > 1]
all.equal(v1_2015$RFACT, fies_2015$RFACT)
```

### 2018

```{r}
keys_2018 <- c("W_REGN", "W_PROV", "SEQUENCE_NO")

nrow(v1_2018[, .N, by = keys_2018][N > 1])
nrow(v2_2018[, .N, by = keys_2018][N > 1])

setkeyv(v1_2018, keys_2018)
setkeyv(v2_2018, keys_2018)
v2_2018_clean <- v2_2018[, !setdiff(intersect(names(v1_2018), names(v2_2018)), keys_2018),
                          with = FALSE]
fies_2018 <- v1_2018[v2_2018_clean]

nrow(v1_2018); nrow(fies_2018)
fies_2018[, .N, by = keys_2018][N > 1]
all.equal(v1_2018$RFACT, fies_2018$RFACT)
```

---

## A.3 Region 10 Datasets and Stacking

```{r}
# Load pre-processed Region 10 files
fies_12 <- read_csv("fies_2012_combined.csv") %>%
  rename(TOTEX = T_TOTEX, FOOD = T_FOOD,
         NFOOD = T_NFOOD, TRANSPORT = T_TRANSPORT) %>%
  mutate(year = 2012)

fies_15 <- read_csv("fies_2015_combined.csv") %>%
  mutate(post = 0, year = 2015)

fies_18 <- read_csv("fies_2018_combined.csv") %>%
  mutate(post = 1, year = 2018)

# Stack all three years
combined <- bind_rows(fies_12, fies_15, fies_18)
```

---

## A.4 Variable Construction

### Treatment Variable

```{r}
# Transport exposure thresholds are defined using 2015 baseline distribution
base_2015 <- combined %>%
  filter(year == 2015)

q1 <- quantile(base_2015$TRANSPORT, 0.10, na.rm = TRUE)  # low-exposure cutoff
q3 <- quantile(base_2015$TRANSPORT, 0.90, na.rm = TRUE)  # high-exposure cutoff

# treat: Binary indicator based on 2015 transport expenditure percentiles
# 1 = high exposure (top 10%); 0 = low exposure (bottom 10%)
# Middle 80% excluded from DiD sample
```

### Post-Intervention Indicator

```{r}
# post: Binary time indicator
# 0 = 2015 (pre-TRAIN Law); 1 = 2018 (post-TRAIN Law)
```

### Outcome Variable

```{r}
# log_totex: Natural log of total household expenditure (TOTEX)
# Log transformation applied to reduce skewness and allow percentage interpretation
```

### Control Variables

```{r}
# FSIZE : Household size
# AGE   : Age of household head
# URB   : Urban/rural indicator (1 = urban, 0 = rural)
# RFACT : Survey expansion weight (used as regression weight)
```

### Construct DiD Dataset

```{r}
did_data <- combined %>%
  filter(year %in% c(2015, 2018)) %>%
  mutate(
    treat = case_when(
      TRANSPORT >= q3 ~ 1,
      TRANSPORT <= q1 ~ 0,
      TRUE            ~ NA_real_
    ),
    post      = if_else(year == 2018, 1, 0),
    log_totex = log(TOTEX)
  ) %>%
  filter(!is.na(treat))

table(did_data$year)
table(did_data$treat, did_data$post)
```

---

## A.5 Difference-in-Differences Model

```{r}
did_model <- lm(
  log_totex ~ treat * post + FSIZE + AGE + URB,
  data    = did_data,
  weights = RFACT
)
summary(did_model)
```

---

## A.6 DiD Plot

```{r}
# Weighted group means
trend_data <- did_data %>%
  filter(!is.na(log_totex), !is.na(treat), !is.na(RFACT)) %>%
  group_by(year, treat) %>%
  summarise(mean_log_totex = weighted.mean(log_totex, w = RFACT, na.rm = TRUE),
            .groups = "drop")

# Counterfactual for treated group
low_2015  <- trend_data %>% filter(year == 2015, treat == 0) %>% pull(mean_log_totex)
low_2018  <- trend_data %>% filter(year == 2018, treat == 0) %>% pull(mean_log_totex)
high_2015 <- trend_data %>% filter(year == 2015, treat == 1) %>% pull(mean_log_totex)
high_2018 <- trend_data %>% filter(year == 2018, treat == 1) %>% pull(mean_log_totex)

cf_high_2018 <- high_2015 + (low_2018 - low_2015)

cf_line <- data.frame(
  year           = c(2015, 2018),
  mean_log_totex = c(high_2015, cf_high_2018)
)
```

```{r, echo=FALSE}
train_did_plot <- ggplot(
  trend_data,
  aes(x = year, y = mean_log_totex,
      color = factor(treat), group = factor(treat))
) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_line(data = cf_line,
            aes(x = year, y = mean_log_totex),
            inherit.aes = FALSE,
            linetype = "dashed", color = "black", linewidth = 1) +
  geom_segment(x = 2018, xend = 2018,
               y = high_2018, yend = cf_high_2018,
               color = "#FF69B4", linewidth = 1.2) +
  scale_color_manual(values = c("0" = "#E07B6A", "1" = "#5B9BD5"),
                     labels = c("Low exposure", "High exposure")) +
  scale_x_continuous(breaks = c(2012, 2015, 2018)) +
  labs(
    title = "Log Total Expenditure Trends of High and Low Exposure Groups",
    x     = "Year",
    y     = "Weighted Mean log Total Expenditure",
    color = "Group"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title         = element_text(face = "bold", size = 15, hjust = 0),
    axis.title         = element_text(size = 15, face = "bold"),
    axis.text          = element_text(size = 15),
    legend.title       = element_text(face = "bold", size = 15),
    legend.text        = element_text(size = 15),
    panel.grid.major   = element_line(color = "gray90", linewidth = 0.4),
    panel.grid.minor   = element_blank(),
    plot.margin        = margin(15, 20, 15, 15)
  )

ggsave("did_plot_train_final.png", plot = train_did_plot,
       width = 8, height = 5, dpi = 300, bg = "white")
train_did_plot
```

---

## A.7 Parallel Trends Assumption Tests

### A.7.1 Pre-Trend Regression Test (2012–2015)

```{r}
# Apply 2015-defined thresholds to 2012 and 2015 data
pretrend_data <- combined %>%
  filter(year %in% c(2012, 2015)) %>%
  mutate(
    treat = case_when(
      TRANSPORT >= q3 ~ 1,
      TRANSPORT <= q1 ~ 0,
      TRUE            ~ NA_real_
    ),
    log_totex = log(TOTEX),
    pre       = if_else(year == 2015, 1, 0)
  ) %>%
  filter(!is.na(treat))

table(pretrend_data$year)
table(pretrend_data$treat, pretrend_data$year)
```

```{r}
pretrend_model <- lm(
  log_totex ~ treat * pre + FSIZE + AGE + URB,
  data    = pretrend_data,
  weights = RFACT
)
summary(pretrend_model)
```

### Pre-Trend Plot

```{r, echo=FALSE}
pretrend_plot_data <- pretrend_data %>%
  group_by(year, treat) %>%
  summarise(mean_log_totex = weighted.mean(log_totex, RFACT, na.rm = TRUE),
            .groups = "drop") %>%
  mutate(group = if_else(treat == 1,
                         "High transport exposure",
                         "Low transport exposure"))

train_pretrend_plot <- ggplot(pretrend_plot_data,
       aes(x = year, y = mean_log_totex,
           color = group, group = group)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = c(2012, 2015)) +
  labs(
    title = "Pre-Trend Check (2012\u20132015)",
    x     = "Year",
    y     = "Weighted Mean of Log Total Expenditure",
    color = "Group"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title       = element_text(face = "bold", size = 15, hjust = 0),
    axis.title       = element_text(size = 15, face = "bold"),
    axis.text        = element_text(size = 15),
    legend.title     = element_text(face = "bold", size = 15),
    legend.text      = element_text(size = 15),
    panel.grid.major = element_line(color = "gray90", linewidth = 0.4),
    panel.grid.minor = element_blank(),
    plot.margin      = margin(15, 20, 15, 15)
  )

ggsave("pretrend_plot_train.png", plot = train_pretrend_plot,
       width = 8, height = 5, dpi = 300, bg = "white")
train_pretrend_plot
```

### A.7.2 Placebo Test

```{r}
# Placebo DiD using only pre-treatment years (2012 vs. 2015)
# A significant interaction here would undermine parallel trends
placebo_model <- lm(
  log_totex ~ treat * pre + FSIZE + AGE + URB,
  data    = pretrend_data,
  weights = RFACT
)
summary(placebo_model)
```

---

## A.8 Event Study

```{r}
# All three survey years; 2015 as reference year (last pre-treatment period)
combined_es <- combined %>%
  filter(year %in% c(2012, 2015, 2018)) %>%
  mutate(
    treat     = case_when(
      TRANSPORT >= q3 ~ 1,
      TRANSPORT <= q1 ~ 0,
      TRUE            ~ NA_real_
    ),
    log_totex = log(TOTEX),
    year      = relevel(factor(year), ref = "2015")
  ) %>%
  filter(!is.na(treat))

event_model <- lm(
  log_totex ~ treat * year + FSIZE + AGE + URB,
  data    = combined_es,
  weights = RFACT
)
summary(event_model)
```

### Event Study Plot

```{r, echo=FALSE}
event_results <- tidy(event_model) %>%
  filter(grepl("treat:year", term)) %>%
  mutate(
    year    = c(2012, 2018),
    ci_low  = estimate - 1.96 * std.error,
    ci_high = estimate + 1.96 * std.error
  )

train_event_study_plot <- ggplot(event_results,
       aes(x = year, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray30") +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high),
                width = 0.4, linewidth = 0.9) +
  geom_point(size = 3) +
  # Reference year point
  annotate("point", x = 2015, y = 0, size = 3, shape = 21,
           fill = "white", color = "gray50", stroke = 1.2) +
  annotate("text", x = 2015, y = -0.005,
           label = "Reference\n(2015)", vjust = 1.4,
           size = 3.5, color = "gray50", fontface = "italic") +
  scale_x_continuous(breaks = c(2012, 2015, 2018)) +
  labs(
    title = "Event Study: Effect of TRAIN Law on Total Expenditure",
    x     = "Year",
    y     = "Treatment Effect on log Total Expenditure"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title       = element_text(face = "bold", size = 15, hjust = 0),
    axis.title       = element_text(size = 15, face = "bold"),
    axis.text        = element_text(size = 15),
    panel.grid.major = element_line(color = "gray90", linewidth = 0.4),
    panel.grid.minor = element_blank(),
    plot.margin      = margin(15, 20, 15, 15)
  )

ggsave("event_study_train.png", plot = train_event_study_plot,
       width = 8, height = 5, dpi = 300, bg = "white")
train_event_study_plot
```



# -------------------------
# OSTER DELTA METHOD
# TRAIN Law: Effect on Log Total Expenditure
# -------------------------

# Step 1: Short regression (no covariates) - beta_dot and R_dot
# Treatment variable is the high/low transport exposure indicator

```{r}
short_model <- lm(
  log_totex ~ treat*post,
  data = did_data,
  weights = RFACT
)
```

# Step 2: Controlled regression (with covariates) - beta_tilde and R_tilde
# Same covariates as did_model, WITHOUT post and interaction
# Oster's method assesses robustness of the treat coefficient
# in the cross-sectional relationship

```{r}
controlled_model <- lm(
  log_totex ~ treat*post + FSIZE + AGE + URB,
  data = did_data,
  weights = RFACT
)
```


# Step 3: Extract values

```{r}
beta_dot   <- coef(short_model)["treat:post"]
beta_tilde <- coef(controlled_model)["treat:post"]
R_dot      <- summary(short_model)$r.squared
R_tilde    <- summary(controlled_model)$r.squared

cat("beta_dot (uncontrolled):", round(beta_dot, 4), "\n")
cat("beta_tilde (controlled):", round(beta_tilde, 4), "\n")
cat("R_dot:", round(R_dot, 4), "\n")
cat("R_tilde:", round(R_tilde, 4), "\n")
```

# -------------------------
# Step 4: Define Rmax values
# Oster (2019) benchmark: Rmax = min($\pi$ * R_tilde, 1)
# -------------------------

# Range of Rmax values to evaluate

```{r}

pi_values  <- seq(1.0, 2.0, by = 0.1)
Rmax_range <- pmin(pi_values * R_tilde, 1)

delta_star_values <- beta_tilde * (R_tilde - R_dot) /
                     ((beta_dot - beta_tilde) * (Rmax_range - R_tilde))

sensitivity_table <- tibble(
  Pi            = pi_values,
  Rmax          = round(Rmax_range, 4),
  Delta_star    = round(delta_star_values, 4),
  Robust        = ifelse(abs(delta_star_values) > 1,
                         "Robust (|δ̃*| > 1)",
                         "Fragile (|δ̃*| ≤ 1)")
)

print(sensitivity_table, n = Inf)
```



```{r}
Rmax_benchmark <- min(1.3 * R_tilde, 1)

cat("\nRmax (1.3 * R_tilde):", round(Rmax_benchmark, 4), "\n")
```

# -------------------------
# Step 5: Oster delta formula (delta_tilde = 1 case)
# Solving for delta that sets beta* = 0:
# delta = [beta_tilde / (beta_dot - beta_tilde)] *
#         [(R_tilde - R_dot) / (Rmax - R_tilde)]
# -------------------------


```{r}
oster_delta <- function(beta_dot, beta_tilde, R_dot, R_tilde, Rmax_benchmark) {
  numerator   <- beta_tilde * (R_tilde - R_dot)
  denominator <- (beta_dot - beta_tilde) * (Rmax_benchmark - R_tilde)
  delta       <- numerator / denominator
  return(delta)
}

delta_benchmark <- oster_delta(beta_dot, beta_tilde,
                               R_dot, R_tilde, Rmax_benchmark)

cat("\n--- Oster Delta Results ---\n")
cat("Delta (Rmax = 1.3 * R_tilde =",
    round(Rmax_benchmark, 4), "):",
    round(delta_benchmark, 4), "\n")
```



# -------------------------
# Step 6: Bias-adjusted beta* (identified set)
# beta* = beta_tilde - delta_tilde *
#         (beta_dot - beta_tilde) * (Rmax - R_tilde) / (R_tilde - R_dot)
# With delta_tilde = 1 (Oster's recommended bound)
# -------------------------


```{r}
bias_adjusted_beta <- function(beta_dot, beta_tilde,
                               R_dot, R_tilde,
                               Rmax_benchmark, delta_tilde = 1) {
  adjustment <- delta_tilde *
    (beta_dot - beta_tilde) *
    (Rmax_benchmark - R_tilde) /
    (R_tilde - R_dot)
  beta_star  <- beta_tilde - adjustment
  return(beta_star)
}

beta_star_benchmark <- bias_adjusted_beta(beta_dot, beta_tilde,
                                          R_dot, R_tilde,
                                          Rmax_benchmark)

cat("\n--- Bias-Adjusted Beta* (delta_tilde = 1) ---\n")
cat("beta* (Rmax = 1.3 * R_tilde):",
    round(beta_star_benchmark, 4), "\n")
```

# -------------------------
# Step 7: Identified set
# [beta_tilde, beta*(Rmax, delta_tilde = 1)]
# -------------------------


```{r}
cat("\n--- Identified Set [beta_tilde, beta*] ---\n")
cat("Rmax = 1.3 * R_tilde: [",
    round(beta_tilde, 4), ",",
    round(beta_star_benchmark, 4), "]\n")
```

# -------------------------
# Step 8: Robustness interpretation
# Robust if identified set excludes zero AND delta >= 1
# -------------------------

```{r}
cat("\n--- Robustness Assessment ---\n")
```


# Benchmark

```{r}
if (delta_benchmark >= 1) {
  cat("Rmax = 1.3*R_tilde: delta =",
      round(delta_benchmark, 4),
      "-- ROBUST (unobservables must be >= as important as",
      "observables to explain away effect)\n")
} else {
  cat("Rmax = 1.3*R_tilde: delta =",
      round(delta_benchmark, 4),
      "-- NOT ROBUST (relatively small unobservable selection",
      "suffices to nullify effect)\n")
}
```


# Check if identified set includes zero

```{r}
sets <- list(
  benchmark = c(beta_tilde, beta_star_benchmark)
)

for (nm in names(sets)) {
  lb <- min(sets[[nm]])
  ub <- max(sets[[nm]])
  if (lb <= 0 & ub >= 0) {
    cat("Identified set (", nm,
        "): INCLUDES zero -- effect not robust",
        "to omitted variable bias\n")
  } else {
    cat("Identified set (", nm,
        "): EXCLUDES zero -- effect survives",
        "omitted variable bias test\n")
  }
}
```




